TacticToe: Learning to Prove with Tactics

### 2018-03-31 tactictoe_jv.tex
presentation
tactictoe-jv: saving orthogonalization
Start editing the paper again

\begin{abstract}
We implement a tactical prover TacticToe on top of the HOL4 interactive theorem
prover. TacticToe learns from human proofs which mathematical technique is
the most suitable in each proof situation.
This knowledge is then used by a Monte Carlo tree search algorithm to
explore the most promising tactic-level proof paths.
On a single CPU, with a time limit of 60 seconds, TacticToe proves 66.4\% of
7164 theorems in the HOL4's standard library whereas
Eprover with auto-schedule solves 34.5\%. The success rate rises to 69.0\% by
combining the results of TacticToe and E~prover.
\end{abstract}

@@

\begin{itemize}
\item Proof recording at the tactic level is made more precise
  (Section~\ref{sec:recording}). We now support pattern
matching constructions and opening SML modules.
%The different
%recording steps (parsing, unfolding and replaying) are regrouped into a single
%\sml function.
\item Monte Carlo tree search (MCTS) replaces A* as our
  search algorithm (Section~\ref{sec:proofsearch}). The MCTS algorithm gives
more balanced feedback on
the success of
each proof step by
comparing subtrees of the search tree instead of leaves. The policy and
evaluation are learned
through supervised learning.
\item Proof guidance required by MCTS is given by prediction algorithms
through supervised learning (Section~\ref{s:prediction}).
These predictors are implemented for three kind of objects: tactics,
theorems and lists of goals.
\item We introduce an orthogonalization process that eliminates redundant tactics (Section~\ref{sec:ortho}).
\item We introduce a tactic abstraction mechanism
  (Section~\ref{sec:synthesis}), which enables us to create
more general and flexible tactics by dynamically
predicting tactic arguments.
\item The internal ATP \metis is complemented
  by asynchronous calls to \eprover (Section~\ref{sec:atp}) to help \tactictoe
during proof search.
\item Evaluation of \tactictoe with a 60 seconds timeout achieves a 66\%
success rate on the standard library. Comparisons between
\tactictoe and \eprover on different type of problems are reported in
Section~\ref{s:experiments}.
\item Minimization and embellishment of the discovered proof facilitates user
  interaction (Section~\ref{sec:proofdisplay}).
\end{itemize}

@@

Thanks to the larger search time available in this experiment, the timeout for
each tactic can be increased from the 0.02 seconds used in \tactictoe's initial
experiments~\cite{tgckju-lpar17} to 0.05
seconds. Removing tactics with similar effect as performed by the
orthogonalization process is only beneficial when running \tactictoe for short
period of time. It seems that the allotted time allows a strategy
running without orthogonalization to catch up.
It is best to predict a list of 16 theorems to instantiate arguments of type
theorem list in tactics. At any rate, argument prediction for abstracted
tactics is the technique that have the highest impact on the success of
\tactictoe. Integration of ATPs being a specialization of this technique
contributes significantly as well. Experiments involving \eprover, run the ATP
using a separate process asynchronously with a timeout of 5 seconds and 128 premises.
The fact that \metis outperforms \eprover in this setting is due to the fact
that \metis is run with a very short timeout, allowing to close different part
of search tree quickly. Because \metis is weaker that \tactictoe as an ATP, it
is given less predictions. But even if an essential lemma for the proof of a
goal is not predicted, a modification on the goal performed by a tactic can
change its prediction to include the necessary lemma.
This effect minimizes the drawback of relying on a small number of predictions.

@@

\begin{table}[h!]
\centering\ra{1.3}
\small
\begin{tabular}{lc}
\toprule
  & Solved (60s) \\
\midrule
   \eprover   & 2472 (34.5\%)\\
   \tactictoe & 4760 (66.4\%)\\
\midrule
   Total  & 4946 (69.0\%)\\
\bottomrule
\end{tabular}
\caption{\label{tab:_param} Evaluation on 7164 top-level theorems of the
\holfour standard library
}
\end{table}

>>>>

\begin{abstract}
We implement a automated tactical prover TacticToe on top of the HOL4 interactive theorem
prover. TacticToe learns from human proofs which mathematical technique is
suitable in each proof situation.
This knowledge is then used in a Monte Carlo tree search algorithm to
explore promising tactic-level proof paths.
On a single CPU, with a time limit of 60 seconds, TacticToe proves 66.4\% of
the 7164 theorems in HOL4's standard library, whereas
E~prover with auto-schedule solves 34.5\%. The success rate rises to 69.0\% by
combining the results of TacticToe and E~prover.
\end{abstract}

@@

\begin{itemize}
%The different
%recording steps (parsing, unfolding and replaying) are regrouped into a single
%\sml function.
\item Monte Carlo tree search (MCTS) replaces A* as our
  search algorithm (Section~\ref{sec:proofsearch}). The MCTS algorithm gives
more balanced feedback on
the success of
each proof step by
comparing subtrees of the search tree instead of leaves. The policy and
value function are learned
through supervised learning.
\item Proof guidance required by MCTS is given by prediction algorithms
through supervised learning (Section~\ref{s:prediction}).
These predictors are implemented for three kinds of objects: tactics,
theorems, and lists of goals.
\item We introduce an orthogonalization process that eliminates redundant tactics (Section~\ref{sec:ortho}).
\item We introduce a tactic abstraction mechanism
  (Section~\ref{sec:synthesis}), which enables us to create
more general and flexible tactics by dynamically
predicting tactic arguments.
\item The internal ATP \metis is complemented
  by asynchronous calls to \eprover (Section~\ref{sec:atp}) to help \tactictoe
during proof search.
\item Evaluation of \tactictoe with a 60 seconds timeout achieves a 66\%
success rate on the standard library. Comparisons between
\tactictoe and \eprover on different type of problems are reported in
Section~\ref{s:experiments}.
\item Proof recording at the tactic level is made more precise
  (Section~\ref{sec:recording}). We now support pattern
matching constructions and opening SML modules.
\item Minimization and embellishment of the discovered proof facilitates user
  interaction (Section~\ref{sec:proofdisplay}).
\end{itemize}

@@

Thanks to the larger search time available in this experiment, the timeout for
each tactic can be increased from the 0.02 seconds used in \tactictoe's initial
experiments~\cite{tgckju-lpar17} to 0.05
seconds. Removing tactics with similar effect as performed by the
orthogonalization process is only beneficial when running \tactictoe for short
period of time. It seems that the allotted time allows a strategy
running without orthogonalization to catch up. Yet, side experiments show
that orthogonalization becomes relevant again when we add additional
tactics through abstraction.
It is best to predict a list of 16 theorems to instantiate arguments of type
theorem list in tactics. At any rate, argument prediction for abstracted
tactics is the technique that have the highest impact on the success of
\tactictoe. Integration of ATPs being a specialization of this technique
contributes significantly as well. Experiments involving \eprover, run the ATP
using a separate process asynchronously with a timeout of 5 seconds and 128 premises.
The fact that \metis outperforms \eprover in this setting is due to the fact
that \metis is run with a very short timeout, allowing to close different part
of search tree quickly. Because \metis is weaker that \tactictoe as an ATP, it
is given less predictions. But even if an essential lemma for the proof of a
goal is not predicted, a modification on the goal performed by a tactic can
change its prediction to include the necessary lemma.
This effect minimizes the drawback of relying on a small number of predictions.

@@

\begin{table}[h!]
\centering\ra{1.3}
\begin{tabular}{lc}
\toprule
  & Solved (60s) \\
\midrule
   \eprover   & 2472 (34.5\%)\\
   \tactictoe & 4760 (66.4\%)\\
\midrule
   Total  & 4946 (69.0\%)\\
\bottomrule
\end{tabular}
\caption{\label{tab:_param} Evaluation on 7164 top-level theorems of the
\holfour standard library
}
\end{table}
